### 2023-06-15 manuscript/appendices/A_fixed_points.qmd
look a kitty !!! *baps with paw*
pwoofwead !!!1

# Fixed point study in the network of neurons

@@

In order to find the fixed points, one has to solve in general $\dot h_i(t) = 0$ for all $i = 1, \cdots, N$. Because of the non-linearity $\phi(h_j(t))$, this is difficult, but we can guess the fixed points $h^\star_i$ and verify them by substituting into the equation that $\dot h_i(t) = 0$.

@@

In the case of the network studied in this work, the find that

@@

-   $h_i^\star = 0 \to h^\star(z_1, \cdots, z_p) = 0$
-   $h_i^\star = F_{\nu i} = z_{\nu i} \to h^\star(z_1, \cdots, z_p) = z_\nu$ for all $\nu = 1, \cdots, p$

[TODO : add derivation for neural field ?]

## Stability of the fixed points

@@

The fixed point is stable if $\vec\delta(t) \to 0$ as $t \to \infty$, in other words if all the eigenvalues of $K$ are negative.

Before we study the spectrum of $K$, we cite the following result : if $\lambda$ is an eigenvalue of $A$, then $\lambda-1$ is an eigenvalue of $A-\mathrm{Id}$. This follows simply from the equation for the eigenvalues $\lambda_i$ and eigenvectors $\vec{v_i}$ of $A$. If $A \vec{v_i} = \lambda_i \vec{v_i}$, then $(A-\mathrm{Id})\vec{v_i}=\lambda_i \vec{v_i} - \vec{v_i} = (\lambda_i - 1)\vec{v_i}$. Therefore to study $K$ it suffices to study the eigenvalues of $J^\prime$.

@@

We study each term separately. Additionally (and similarly as before), we make the assumption that the functions $g_\mu$ and the functions $f_\mu$ are component-wise, and that the distribution $\rho$ factorises.

@@

We now consider a Taylor expansion of the rate function $\phi(h)$ up to order 3, and cite results obtained from symbolic computation and known moments of the gaussian distribution.

@@

* When $\phi$ is linear, we see that all the eigenvalues are zero, therefore the first order stability is unconclusive. Numerically, we see that when $\phi$ is linear, the convergence to the fixed points is very slow.
* When $\phi$ is quadratic, the zero fixed point is stable, and the pattern fixed point is unstable, regardless of the sign of the coefficients.
* When $\phi$ is cubic, the sign of the eigenvalues depends primarily on $c_3$ (negative $c_1$ breaks the assumption that $\phi$ must be monotonically increasing). The intuition is that we require $c_1$ to be large enough (in other words, the growth of the firing rate when $h$ becomes positive) and $c_3 < 0$ such that $3 c_1 c_3 + 15 c_3^2 < 0$.
  The magnitude of the second order term $c_2$ only contributes to adding instability to the pattern fixed point, and stability to the zero fixed point, which leads to the conclusion it is best set to $c_2=0$ if we want the network to be stable at the pattern fixed points.

@@

In conclusion, the spectrum of $K$ is composed of $p$ eigenvalues $\lambda_\nu = \avg{g_\nu(z_\nu) f_\nu(z_\nu) \partial \phi(h^\star)}-1$, and $N-p$ eigenvalues $\lambda_\perp = -1$. For the network considered in this work, the interpretation of this is that components of the field aligned with the patterns $z_\nu$ converge to the fixed point, and the orthogonal components decay as $t \to \infty$. The fixed point $h_i^\star=0$ is therefore unstable, and the pattern fixed points $h_i^\star = z_{\nu i}$ are stable. We also note that since the normal distribution is centered around zero and symmetric, $h_i^\star = -z_{\nu i}$ are also stable fixed points.

>>>>

# Fixed point study in the network of neurons {#sec-fixedpoints}

@@

To find the fixed points, one has to solve in general $\dot h_i(t) = 0$ for all $i = 1, \cdots, N$. Because of the non-linearity $\phi(h_j(t))$, this is difficult, but we can guess the fixed points $h^\star_i$ and verify them by substituting into the equation that $\dot h_i(t) = 0$.

@@

In the case of the network studied in this work, we find that

@@

* $h_i^\star = 0 \to h^\star(z_1, \cdots, z_p) = 0$
* $h_i^\star = F_{\nu i} = z_{\nu i} \to h^\star(z_1, \cdots, z_p) = z_\nu$ for all $\nu = 1, \cdots, p$

## Stability of the fixed points {#sec-fixedpoints-stability}

@@

The fixed point is stable if $\vec\delta(t) \to 0$ as $t \to \infty$, in other words, if all the eigenvalues of $K$ are negative.

Before we study the spectrum of $K$, we cite the following result: if $\lambda$ is an eigenvalue of $A$, then $\lambda-1$ is an eigenvalue of $A-\mathrm{Id}$. This follows simply from the equation for the eigenvalues $\lambda_i$ and eigenvectors $\vec{v_i}$ of $A$. If $A \vec{v_i} = \lambda_i \vec{v_i}$, then $(A-\mathrm{Id})\vec{v_i}=\lambda_i \vec{v_i} - \vec{v_i} = (\lambda_i - 1)\vec{v_i}$. Therefore to study $K$ it suffices to study the eigenvalues of $J^\prime$.

@@

We study each term separately. Additionally (and similarly as before), we assume that the functions $g_\mu$ and the functions $f_\mu$ are component-wise, and that the distribution $\rho$ factorises.

@@

We now consider a Taylor expansion of the rate function $\phi(h)$ up to order 3, and cite results obtained from symbolic computation and known moments of the Gaussian distribution.

@@

* When $\phi$ is linear, we see that all the eigenvalues are zero, therefore the first order stability is inconclusive. Numerically, we see that when $\phi$ is linear, the convergence to the fixed points is very slow.
* When $\phi$ is quadratic, the zero fixed point is stable, and the pattern fixed point is unstable, regardless of the sign of the coefficients.
* When $\phi$ is cubic, the sign of the eigenvalues depends primarily on $c_3$ (negative $c_1$ breaks the assumption that $\phi$ must be monotonically increasing). The intuition is that we require $c_1$ to be large enough (in other words, the growth of the firing rate when $h$ becomes positive) and $c_3 < 0$ such that $3 c_1 c_3 + 15 c_3^2 < 0$.
  The intuition behind this is that if the slope $c_1$ is stronger, then the bump of $\partial \phi$ becomes "more concentrated", which helps lower the eigenvalues associated with the pattern fixed points.
  The magnitude of the second order term $c_2$ only contributes to adding instability to the pattern fixed point, and stability to the zero fixed point, which leads to the conclusion it is best set to $c_2=0$ if we want the network to be stable at the pattern fixed points.

@@

In conclusion, the spectrum of $K$ is composed of $p$ eigenvalues $\lambda_\nu = \avg{g_\nu(z_\nu) f_\nu(z_\nu) \partial \phi(h^\star)}-1$, and $N-p$ eigenvalues $\lambda_\perp = -1$. For the network considered in this work, the interpretation of this is that components of the field aligned with the patterns $z_\nu$ converge to the fixed point, and the orthogonal components decay as $t \to \infty$. The fixed point $h_i^\star=0$ is therefore unstable, and the pattern fixed points $h_i^\star = z_{\nu i}$ are stable. We also note that since the normal distribution is centred around zero and symmetric, $h_i^\star = -z_{\nu i}$ are also stable fixed points.

### 2023-06-15 manuscript/chapters/1_intro.qmd
pwoofwead !!!1

As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual [@SatNau12;@MulRey14], somatosensory [@PetHah03;@FerBol06], motor [@RubRob06;@TakKim15], and hippocampal [@LubSia09;@PatFuj12;@PatSch13] cortices (see @MulCha18 for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings @Bre17. See also @DipRan18 for an example of neural field modelling of 2-photon calcium imaging recording of the visual cortex.

@@

The goal of this Master Thesis is to show that the answer to the second question is highly nontrivial because even the dimensionality of the embedding space, for some given spatiotemporal dynamics, is not clearly defined. This means that it is possible to find two embeddings (and connectivity kernels) of different dimensions that give rise to identical dynamics; contradicting the intuition that the embedding dimension should at least be as high as the dimensionality of the dynamics.

In particular, we show that for any two-dimensional neural field equation (that is, with two spatial dimensions), there is a one-dimensional neural field equation on the interval, where the embedding space is simply the interval $[0,1]$, from which the solution to the two-dimensional equation can be fully reconstructed. The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections for the square $[0,1]^2$ and the interval $[0,1]$. Importantly, these bijections are not (and can't be) diffeomorphisms. Also, we show that although the connectivity kernel of the one-dimensional equation seems to have a fractal structure, it is sufficiently regular for its solution to be numerically approximated using standard mesh-based simulations.

@@

If the embedding is well-chosen,[^1] the connectivity matrix $J_{ij}$ approaches a connectivity kernel $w(\vec y, \vec z)$ as $N \rightarrow \infty$, and neurons close in the embedding space tend to have similar potentials, such that we can write the continuous *neural field* $h(t, \vec z)$ of neuron potentials. Before we write down the neural field equation which describes how the field evolves in time, we need to introduce the neuron density $\rho(\vec z)$, which measures the number of neurons per unit volume of the embedding space. The neural field equation is then written as such:

@@

$$
\partial_t h(t, \vec z) = -h(t, \vec z) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(t, \vec y)) \rho(\mathrm d \vec y).
$$ {#eq-nf}

@@

$$
\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \left[\lambda \circ S^{-1}\right](\d \beta)
$$

@@

@sec-nf introduces a toy model for a neural field in $p$ dimensions, and we discuss interpretations of numerical simulation schemes.

@sec-mappings is dedicated to the discussion of mappings from 2 dimensions to 1 dimension, we give intuitions accompanied by simulations, and discuss in detail how "locality" emerges from these mappings.

>>>>

As models of spatiotemporal neural population dynamics on the cortical sheet, neural field models can be used to explain experimentally observed patterns of cortical activity, such as travelling waves in visual [@SatNau12;@MulRey14], somatosensory [@PetHah03;@FerBol06], motor [@RubRob06;@TakKim15], and hippocampal [@LubSia09;@PatFuj12;@PatSch13] cortices (see @MulCha18 for a review), and are used to model large-scale brain signals, such as electroencephalography (EEG) recordings @Bre17. See also @DipRan18 for an example of neural field modelling of calcium imaging recording of the visual cortex.

@@

The goal of this Master Thesis is to show that the answer to the second question is highly nontrivial because even the dimensionality of the embedding space, for some given spatiotemporal dynamics, is not clearly defined. This means that it is possible to find two embeddings (and connectivity kernels) of different dimensions that give rise to identical dynamics; *contradicting the intuition that the embedding dimension should at least be as high as the dimensionality of the dynamics* (the number of variables needed to fully describe the dynamical system).

In particular, we show that for any two-dimensional neural field equation (that is, with two spatial dimensions), there is a one-dimensional neural field equation on the interval, where the embedding space is simply the interval $[0,1]$, from which the solution to the two-dimensional equation can be fully reconstructed. The mapping from the two-dimensional equation to the one-dimensional equation is done via well-known measurable bijections between the square $[0,1]^2$ and the interval $[0,1]$. Importantly, these bijections are not (and can't be) diffeomorphisms. Also, we show that although the connectivity kernel of the one-dimensional equation seems to have a fractal structure, it is sufficiently regular for its solution to be numerically approximated using standard mesh-based simulations.

@@

If the embedding is well-chosen,[^1] the connectivity matrix $J_{ij}$ approaches a connectivity kernel $w(\vec y, \vec z)$ as $N \rightarrow \infty$, and neurons close in the embedding space tend to have similar potentials, such that we can write the continuous *neural field* $h(\vec z, t)$ of neuron potentials. Before we write down the neural field equation which describes how the field evolves in time, we need to introduce the neuron density $\rho(\vec z)$, which measures the number of neurons per unit volume of the embedding space. The neural field equation is then written as such:

@@

$$
\partial_t h(\vec z, t) = -h(\vec z, t) + \int_{\mathbb{R}^p} w(\vec z, \vec y) \phi(h(\vec y, t)) \rho(\mathrm d \vec y).
$$ {#eq-nf}

@@

$$
\partial_t \tilde h(\alpha, t) = -\tilde h(\alpha, t) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta)
$$

@@

@sec-nf introduces a toy model for a neural field in $p$ dimensions, and tools to study its dynamics. We also discuss interpretations of numerical simulation schemes.

@sec-mappings is dedicated to the discussion of mappings from 2 dimensions to 1 dimension. We give intuitions accompanied by simulations, and discuss in detail how "locality" emerges from these mappings.

### 2023-06-15 manuscript/chapters/2_neural_fields.qmd
pwoofwead !!!1

$$
G_{\mu i} = \tilde\phi(z_{i \mu}) \stackrel{\text{def}}{=} \frac{\phi(z_{i \mu}) - \avg{z_{i \mu}}}{\mathrm{Var}[\phi(z_{i \mu})]}.
$$

@@

* $h^\star_i = z_{\mu i}$ is a stable fixed point for all $\mu \in \{1,\cdots,p\}$. Additionally, because the Gaussian distribution is symmetric around zero, $h_i = -z_{\mu i}$ are also stable fixed points. We refer to these fixed points as the "pattern" fixed points, because the network of neurons "remembers" the vectors $\vec{F_1}, \cdots, \vec{F_p}$.
* $h^\star_i=0$ is an unstable fixed point. It corresponds to all the neuron potentials being set to zero.

@@

1. Taking Taylor expansions of the activation functions reveals that -- at least up to order 3 -- uneven powers tend to increase stability of the pattern fixed points (and reciprocally, decrease stability of the zero fixed point), and vice-versa for the even powers (with the notable exception of the constant offset, which doesn't play a role in this analysis).
2. A steeper slope at the origin of the activation function also seems to improve stability of the pattern fixed points. This corresponds to a sharper difference between the "inactive" (low potential) and the "active" (high potential) neurons. The intuition behind this is that if the slope is stronger, then the bump of $\partial \phi$ becomes "more concentrated", which helps lower the eigenvalues associated to the pattern fixed points.
3. The spectrum of $K$ is composed of $p$ eigenvalues that depend on the fixed point, and $N-p$ eigenvalues $\lambda_\perp = -1$ corresponding to the orthogonal complement of the pattern fixed points.

@@

We should however note that the results obtained in @sec-fixedpoints are valid in the $N \to \infty$ limit. A numerical analysis shows in @fig-spectrum that already at $N=64$, the analytic eigenvalues match the numerical eigenvalues closely, and this correspondence improves when we take larger $N$ (typically we will for the rest of this Thesis take $N > 10^3$).

@@

We first demonstrate what happens with numerical simulations in the simple case of $p=1$. We generate a network of 50,000 neurons, where each neuron samples a one-dimensional Gaussian to get its position. The function $\tilde \phi$ is computed by rescaling $\phi$ with the numerically computed mean and variance.[^code1]

@@

We however note that in this case, there are two pattern fixed points that the network can converge to. In the example given here the network coincidentally, due to finite numerical effects, converged again to the first pattern fixed point, which we can see because the neural field is constant in the $z_2$ direction. We write the fixed point $h^\star(z_1, z_2, t) = z_1$. The neural field is now written as

@@

$$
\begin{aligned}
\mathcal{I}(\vec z) &= \int_{\Rp} \underbrace{w(\vec z, \vec y) \phi(h(t, \vec y))}_{I(\vec z, \vec y)} \rho(\d \vec y) \\
\hat{\mathcal{I}}(\vec z) &= \frac 1N \sum_{i=1}^N I(\vec z, \vec{y_i}), \, \vec{y_i} \stackrel{\text{iid}}{\sim} \rho
\end{aligned}
$$ {#eq-mc}

@@

One small nuance we should add is that strictly speaking, Monte-Carlo integration should resample the integrand at every timestep, but the network of neurons samples the neurons once, and propagates these points through time.

@@

Another method to estimate an integral numerically is by regularly sampling the integrand on a fixed grid. We simply state the result that the convergence rate of the estimation is $\bO(N^{-2/p})$. Interestingly, the convergence rate depends on the dimension $p$ of the integral. For $p \geq 4$, Monte-Carlo integration shows better scaling behaviour (but depending on the use case, prefactors of the error can change which method is better).

@@

We now introduce some tools that will help us to understand the toy model. We stress that in general, other neural field models might not have this luxury, and that we only make use of these tools to help explain the dynamics of the neural fields. Computationally, the overlaps and projections introduced here are only computed from the results of the simulation (see [`overlap.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_overlap.py) in the source code).

@@

$$
\begin{aligned}
m_\mu(t) &= \frac 1N \sum_{i=1}^N \tilde \phi(z_{\mu i}) \phi(h_i(t)) \\
&\xrightarrow{N \to \infty} \int_{\Rp} \tilde \phi(z_\mu) \phi(h(\vec z, t)) \rho(\d \vec z)
\end{aligned}
$$ {#eq-overlap}

@@

This formulation "hides" the integral inside the overlap variables, but one should not forget that the overlaps depend on the neural field at a given time.

@@

$$
\kappa_\mu(t) = \int_{\Rp} y_\mu h(\vec y, t) \rho(\d \vec y),
$$ {#eq-kappa}

@@

$$
\begin{aligned}
\dot \kappa_\mu(t) &= -\kappa_\mu(t) + \int_{\Rp} \tilde\phi(y_\mu) \phi(h(\vec y, t)) \rho(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t) \\
\kappa_\mu(0) &= \int_{\mathbb{R}^p} y_\mu h(\vec y, 0) \rho(\mathrm d \vec y),
\end{aligned}
$$

@@

Additionally, it can be seen from the expression of the overlaps in @eq-overlap, that in the case of a linear activation function $\phi(h) = c_0 + c_1 h$, then $m_\mu(t) = \kappa_\mu(t)$.

## A cycling neural field

@@

We note however that @eq-nf-cycling-p no longer is a partial differential equation, but rather a delayed differential equation, which are in general much more complicated to solve. A simplifying assumption that we make is that we extend the initial condition back in time, such that $h(\vec z, t < 0) = h(\vec z, 0)$ is a constant function. Numerically, this also means that we have to store a history of the neural fields instead of only the current one. Since the `solve_ivp` method dynamically adapts the timestep of integration, we use linear interpolation between two known points in the stored neural field states to estimate $h(\vec z, t-\delta)$ (see [`lagging.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_lagging.py) and [its use](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L334)).

With these considerations, we simulate a $p=2$ cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$ in @fig-cycling-anim. The latent trajectory is estimated from the simulation results, and animated simultaneously to the neural field. We clearly see the oscillations between the patterns $\mu=1$ and $\mu=2$.

@@

In preparation for mapping of $[0,1]^2$ to $[0,1]$, we first map the 2-dimensional neural field equation in $\R^2$ to $[0,1]^2$, and, in general $\Rp$ to $[0,1]^p$. There are many ways to do this by using functions $\R \mapsto [0,1]$, but a practical choice is the Gaussian CDF (Cumulative Density Function), because it has the benefit of absorbing the density in the kernel, in the sense that the integral $[0,1]^p$ becomes weighted by the uniform distribution. The one-dimensional Gaussian CDF is defined as

@@

We make the final remark that this method of mapping $\Rp$ to $[0,1]^p$ does not work for any neural field. In our case, we use the fact that the distribution factorizes into $\rho(z_1, \cdots, z_p) = \rho_1(z_1) \cdots \rho_p(z_p)$, such that the components of $\vec{v}$ are sampled independently. We can then define the "componentwise CDF" that relates to the total CDF via

>>>>

$$
G_{\mu i} = \tilde\phi(z_{\mu i}) \stackrel{\text{def}}{=} \frac{\phi(z_{\mu i}) - \avg{z_{\mu i}}}{\mathrm{Var}[\phi(z_{\mu i})]}.
$$

@@

* $h^\star_i = z_{\mu i}$ is a stable fixed point for all $\mu \in \{1,\cdots,p\}$. Additionally, because the Gaussian distribution is symmetric around zero, $h^\star_i = -z_{\mu i}$ are also stable fixed points. We refer to these fixed points as the "pattern" fixed points, because the network of neurons "remembers" the vectors $\vec{F_1}, \cdots, \vec{F_p}$.
* $h^\star_i=0$ is an unstable fixed point. It corresponds to all the neuron potentials being set to zero.

@@

1. Taking Taylor expansions of the activation functions reveals that -- at least up to order 3 -- uneven powers tend to increase stability of the pattern fixed points (and reciprocally, decrease stability of the zero fixed point), and vice-versa for the even powers (with the notable exception of the constant offset, which doesn't play a role in this analysis).
2. A steeper slope at the origin of the activation function also seems to improve stability of the pattern fixed points. This corresponds to a sharper difference between the "inactive" (low potential) and the "active" (high potential) neurons.
3. The spectrum of $K$ is composed of $p$ eigenvalues that depend on the fixed point, and $N-p$ eigenvalues $\lambda_\perp = -1$ corresponding to the orthogonal complement of the pattern fixed points.

@@

We should however note that the results obtained in @sec-fixedpoints are valid in the $N \to \infty$ limit. A numerical analysis shows in @fig-spectrum that already at $N=64$, the analytic eigenvalues match the numerical eigenvalues closely, and this correspondence improves when we take larger $N$ (typically we will for the rest of this thesis take $N > 10^3$).

@@

We first demonstrate what happens with numerical simulations in the simple case of $p=1$. We generate a network of 50,000 neurons, where each neuron samples a one-dimensional Gaussian to get its position in the embedding. The function $\tilde \phi$ is computed by rescaling $\phi$ with the numerically computed mean and variance.[^code1]

@@

We however note that in this case, there are two pattern fixed points that the network can converge to. In the example given here, the network coincidentally converged again to the first pattern fixed point. The network is in the $\mu=1$ pattern, which we can see because the neural field is constant in the $z_2$ direction. We write the fixed point $h^\star(z_1, z_2, t) = z_1$. The neural field is now written as

@@

$$
\begin{aligned}
\mathcal{I}(\vec z) &= \int_{\Rp} \underbrace{w(\vec z, \vec y) \phi(h(\vec y, t))}_{I(\vec z, \vec y)} \rho(\d \vec y) \\
\hat{\mathcal{I}}(\vec z) &= \frac 1N \sum_{i=1}^N I(\vec z, \vec{y_i}), \, \vec{y_i} \stackrel{\text{iid}}{\sim} \rho
\end{aligned}
$$ {#eq-mc}

@@

One small nuance we should add, is that strictly speaking, Monte-Carlo integration should resample the integrand at every timestep, but the network of neurons samples the neurons once, and propagates these points through time.

@@

Another method to estimate an integral numerically is by discretizing the integrand on a regular grid. We simply state the well-known result that the convergence rate of the estimation is $\bO(N^{-2/p})$. Interestingly, the convergence rate depends on the dimension $p$ of the integral. For $p \geq 4$, Monte-Carlo integration shows better scaling behaviour (but depending on the use case, prefactors of the error can change which method is better).

@@

We now introduce some tools that will help us understand the toy model. We stress that in general, other neural field models might not have this luxury, and that we only make use of these tools to help explain the dynamics of the simulated neural fields. The overlaps and projections introduced here are only computed from the results of the simulation (see [`overlap.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_overlap.py) in the source code).

@@

$$
\begin{aligned}
m_\mu(t) &= \frac 1N \sum_{i=1}^N \tilde \phi(z_{\mu i}) \phi(h_i(t)) \\
&\xrightarrow{N \to \infty} \int_{\Rp} \tilde \phi(z_\mu) \phi(h(\vec z, t)) \mathcal{N}^p(\d \vec z)
\end{aligned}
$$ {#eq-overlap}

@@

This formulation "hides" the integral inside the overlap variables, but one should not forget that the overlaps depend on the neural field at any given time.

@@

$$
\kappa_\mu(t) = \int_{\Rp} y_\mu h(\vec y, t) \mathcal{N}^p(\d \vec y),
$$ {#eq-kappa}

@@

$$
\begin{aligned}
\dot \kappa_\mu(t) &= -\kappa_\mu(t) + \int_{\Rp} \tilde\phi(y_\mu) \phi(h(\vec y, t)) \mathcal{N}^p(\mathrm d \vec y) = -\kappa_\mu(t) + m_\mu(t) \\
\kappa_\mu(0) &= \int_{\mathbb{R}^p} y_\mu h(\vec y, 0) \mathcal{N}^p(\mathrm d \vec y),
\end{aligned}
$$

@@

Additionally, it can be seen from the expression of the overlaps in @eq-overlap, that in the case of a linear activation function $\phi(h) = c_0 + c_1 h$, we have the equality $m_\mu(t) = \kappa_\mu(t)$.

## A cycling neural field {#sec-cyling-nf}

@@

We note however that @eq-nf-cycling-p no longer is a partial differential equation, but rather a delayed differential equation, which are in general much more complicated to solve. A simplifying assumption that we make is we extend the initial condition back in time, such that $h(\vec z, t < 0) = h(\vec z, 0)$ is a constant function. Numerically, this also means that we have to store a history of the neural fields instead of only the current one. Since the [`solve_ivp`](https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.solve_ivp.html) method dynamically adapts the timestep of integration, we use linear interpolation between two known points in the stored neural field states to estimate $h(\vec z, t-\delta)$ (see [`lagging.py`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_lagging.py) and [its use](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L334)).

With these considerations, we simulate a $p=2$ cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$ in @fig-cycling-anim. The latent trajectory is estimated from the simulation results, and animated simultaneously with the neural field. We clearly see the oscillations between the patterns $\mu=1$ and $\mu=2$.

@@

In preparation to the next chapter, we first map the 2-dimensional neural field equation in $\R^2$ to $[0,1]^2$, and, in general $\Rp$ to $[0,1]^p$. There are many ways to do this by using functions $\R \mapsto [0,1]$, but a practical choice is the Gaussian CDF (Cumulative Density Function), because it has the benefit of absorbing the density into the kernel, in the sense that the integral $[0,1]^p$ becomes weighted by the uniform distribution. The one-dimensional Gaussian CDF is defined as

@@

We make the final remark that this method of mapping $\Rp$ to $[0,1]^p$ by using the CDF does not work for any neural field. In our case, we use the fact that the distribution factorizes into $\rho(z_1, \cdots, z_p) = \rho_1(z_1) \cdots \rho_p(z_p)$, such that the components of $\vec{v}$ are independent. We can then define the "componentwise CDF" that relates to the total CDF via
